Embedded Scalars: {A}n Experimental Revisit

### 2014-05-21 text/paper/paper4.tex
statistical analyses critical

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses using log-linear
models (i.e., generalized linear regression for count data). To
complement the picture, we also provide a Bayesian analysis that
probes specifically into the preference relations between candidate
readings.

@@

\caption{Positional judgments for the four target conditions. Bars
  show the total number of \emph{yes} (up) and \emph{no} judgments
  (down) against the position within a sequence. E.g., we recorded a
  total of 103 \emph{yes} judgments at the third position in the
  \as-sequence with neutral intonation, which corresponds to 103
  answers that were classified as \emph{literal} answers. Notice that
  the \as-condition has seven positions, the \es-condition only five.}
\label{fig:JudgmentsK2}
\end{figure}
%
We coded these answers as {\it literal}, {\it global} or {\it local}
if they were as expected under one of these readings and as {\it
  error} if not. The distribution of readings thus coded is presented
in Table~\ref{tab:CC-ContingencyTable}. The table also classifies
responses according to the position (1--3) in the experiment, i.e.,
whether it was the first, second or third time that a participant saw
the critical condition in question.
%
\begin{table}[h]
\centering

@@

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are
7.02\% / 4.39\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.30\% / 17.54\% 'error' in the neutral /
accented \es-conditions. A look at Figure~\ref{fig:JudgmentsK2}
reveals that there is some systematicity in 'error' answers. Most
answers that we classified as 'errors' are plausibly explained as a
few random errors, a handful of mistakes resulting from confusing
buttons for \emph{true} and \emph{false} and a few 'spill-overs' where
participants gave responses one position too late. There is one
notable exception, though. A substantial number of \emph{true}
responses where given at the second position of \es-sequences (9.65\%
/ 8.77\%) which is shown in Figure~\ref{fig:exseqES2}. Although this
position in the sequence is the critical position to reveal a global
reading, the latter would be indicated by a \emph{false} response. It
appears that subjects read \es-sentences as if the non-monotonic
quantifier \emph{exactly one} was a non-monotonic \emph{at least one},
without pragmatic enrichment of \emph{some}. 16 out of the total 21
answers in this category come from 5 subjects, while the remaining 5
were single answers by 5 different subjects. \todo{subjects 1, 9, 12,
  21, 84} This suggests that at least for some subjects non-monotonic
quantifiers were problematic. \todo{how to deal with this in the future?}

We fitted generalized linear models to the contingency table data in
Table~\ref{tab:CC-ContingencyTable}, using factors \textsc{Condition}
with levels \as and \es, \textsc{Accent} with levels \emph{neutral}
and \emph{accented}, and \textsc{Position} with levels 1 through 3 to
predict the dependent factor \emph{Response} with levels
\emph{literal}, \emph{local}, and \emph{other}. The latter levels
lumps \emph{global} and \emph{error} responses together, because
otherwise we would not have enough cell counts to apply log-linear
regression. We selected the ``best'' predictor of the data by a
gradient search over hierarchical models in terms of AICs, starting
from the saturated model. The best model takes main factors
\textsc{Response} and \textsc{Condition} into account, as well as the
interaction term \textsc{Response:Condition} ($\chi^2 = 17.19$,
$df=42$, $p = 0.99$, $AIC = 158.01$ (compared to $AIC=208.25$ of the
saturated model)). Crucially, factors \textsc{Position} and
\emph{Accent} are spurious in the best model. This suggests that
response patterns only depend on the type of the sentence, but that
there were no learning or fatigation effects and that accentuation did
not influence answer patterns significantly.

In order to test whether there is support for the postulation of local
readings, we compared the previous ``best'' model to a model that only
differs from the former in that it subsumes the level \emph{local} of
factor \textsc{Reading} under level \emph{other} as well
\citep[e.g.][Chapter~15]{Crawley2007:The-R-Book}. Model comparison
reveals that there is no significant improvement in explanatory power
($Pr(|\chi|) = 0.2689$) of the more complex model that includes the
level \emph{local} (df=30, residual deviance = 9.76) over the leaner
model without this level (df=32, residual deviance = 12.39,
AIC=156.63). This means that our data does not provide strong evidence
for the maintenance of belief in \emph{local} readings in our
setting. (Subsequent analysis will shed more light on this.)

\paragraph{Preference-related controls.} Positional answers for our
preference-related control conditions are shown in
Figure~\ref{Fig:TF_judgments}.
%
\begin{figure}[]
\centering

@@

\caption{Positional answers for \lc-sequences and
  \ec-sequences. \emph{True} judgements are plotted upwards on each
  position, \emph{false} judgements downwards.}
\label{Fig:TF_judgments}
\end{figure}
%
Judgments were coded as \ec or as \lc if they were as expected under
one of these readings, and as \emph{error} if they were not. The
distributions of readings thus coded, for each position at which this
critical item occurred during the experiment, is given in
Table~\ref{tab:CC-ContingencyTable-TF}.

\begin{table}[]
\centering

\begin{tabular}{lcccc}
  condition & position & \multicolumn{3}{c}{response type} \\ \cmidrule(r){3-5}
  && \emph{LC} & \emph{EC} & \emph{error}  \\
  \midrule \addlinespace[0.15cm]
  \textbf{\lc neutral} 
  & Pos 1 & 27 &  7  & 4 \\
  & Pos 2 & 23 &  11 & 4 \\
  & Pos 3 & 26 &  9  & 3 \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{103} &  \textbf{3} & \textbf{0}  \\
  \addlinespace[0.25cm]
  \textbf{\lc with \lc cue} 
  & Pos 1 & 26 &  6 & 5  \\
  & Pos 2 & 31 &  5 & 2  \\
  & Pos 3 & 33 &  3 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2}  \\
  \addlinespace[0.25cm]
  \textbf{\lc with \ec cue} 
  & Pos 1 & 14 &  18 & 3  \\
  & Pos 2 & 16 &  16 & 1  \\
  & Pos 3 & 14 &  17 & 5  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2}  \\
  \addlinespace[0.25cm]
  \textbf{\ec neutral} 
  & Pos 1 & 22 &  6 & 10  \\
  & Pos 2 & 28 &  6 & 4  \\
  & Pos 3 & 28 &  8 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2}  \\
  \addlinespace[0.25cm]
  \textbf{\ec with \lc cue} 
  & Pos 1 & 30 &  5 & 3  \\
  & Pos 2 & 32 &  5 & 1  \\
  & Pos 3 & 30 &  3 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2}  \\
  \addlinespace[0.25cm]
  \textbf{\ec with \ec cue} 
  & Pos 1 & 19 &  10 & 9  \\
  & Pos 2 & 19 &  7  & 12  \\
  & Pos 3 & 19 &  7  & 12  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2}  \\
\end{tabular}

@@

>>>>

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses using log-linear
models (i.e., generalized linear regression for count data). To
complement the picture, we also report a Bayesian analysis that probes
specifically into the preference relations between candidate readings.

@@

\caption{Answers for the \as- and \es-sequences. Bars show the total
  number of \emph{yes} (up) and \emph{no} judgments (down) collected
  at each step during a sequence. E.g., we recorded a total of 103
  \emph{yes} judgments at the third position in the \as-sequence with
  neutral intonation, which corresponds to 103 answers that were
  classified as \emph{literal} answers. Notice that the \as-condition
  had seven steps, the \es-condition only five.}
\label{fig:JudgmentsK2}
\end{figure}
%
We coded these answers as {\it literal}, {\it global} or {\it local}
if they were as expected under one of these readings and as {\it
  error} if not. The distribution of readings thus coded is given in
Table~\ref{tab:CC-ContingencyTable}. The table also classifies
responses according to the position (1--3) in the experiment, i.e.,
whether it was the first, second or third time that a participant saw
the critical condition in question.
%
\begin{table}[h]
\centering

@@

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are
7.02\% / 4.39\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.30\% / 17.54\% in the neutral / accented
\es-conditions. A look at Figure~\ref{fig:JudgmentsK2} reveals that
there is some systematicity in 'error' answers. Most answers that we
classified as 'errors' are plausibly explained as a few random errors,
a handful of mistakes resulting from confusing buttons for \emph{true}
and \emph{false} and a few 'spill-overs' where participants gave
responses one position too late. There is one notable exception,
though. A substantial number of \emph{true} responses where given at
the second position of \es-sequences (9.65\% / 8.77\%) which is shown
in Figure~\ref{fig:exseqES2}. Although this position in the sequence
is the critical position to reveal a global reading, the latter would
be indicated by a \emph{false} response. It appears that subjects read
\es-sentences as if the non-monotonic quantifier \emph{exactly one}
was a monotonic \emph{at least one}, without pragmatic enrichment of
\emph{some}.\fn{If \emph{some} is pragmatically strengthened under the
  scope of \emph{at least one}, we would expect \emph{true} judgements
  at position 3. Although exactly 3 \emph{true} answers occurred there
  for each neutral and accented \es-conditions, the main point to
  notice is that this misreading of the non-monotonic quantifier does
  not explain \emph{true} judgements at position 5. These are
  therefore still indicative of local readings.} 16 out of the total
21 answers in this category come from 5 subjects, while the remaining
5 were single answers by 5 different subjects. \todo{subjects 1, 9,
  12, 21, 84} This suggests that at least for some subjects
non-monotonic quantifiers were problematic. \todo{how to deal with
  this in the future?}

We fitted generalized linear models to the contingency table data in
Table~\ref{tab:CC-ContingencyTable}, using factors \textsc{Condition}
with levels \as and \es, \textsc{Accent} with levels \emph{neutral}
and \emph{accented}, and \textsc{Position} with levels 1 through 3 to
predict the dependent factor \textsc{Reading} with levels
\emph{literal}, \emph{local}, and \emph{other}. The latter levels
lumps \emph{global} and \emph{error} responses together, because
otherwise we would not have enough cell counts to apply log-linear
regression. We selected the ``best'' predictor of the data by a
gradient search over hierarchical models in terms of AICs, starting
from the saturated model. The best model takes main factors
\textsc{Response} and \textsc{Condition} into account, as well as the
interaction term \textsc{Response:Condition} ($\chi^2 = 17.19$,
$df=42$, $p = 0.99$, $AIC = 158.01$ (compared to $AIC=208.25$ of the
saturated model)). Crucially, factors \textsc{Position} and
\emph{Accent} are spurious in the best model. This suggests that
response patterns only depend on the type of the sentence, but that
there were no learning or fatigation effects and that accentuation did
not influence answer patterns significantly.

In order to test whether there is support for the postulation of local
readings, we compared the previous ``best'' predictor to a model that
only differs from the former in that it subsumes the level
\emph{local} of factor \textsc{Reading} under level \emph{other} as
well \citep[e.g.][Chapter~15]{Crawley2007:The-R-Book}. Model
comparison reveals that there is no significant improvement in
explanatory power ($Pr(|\chi|) = 0.2689$) of the more complex model
that includes the level \emph{local} (df=30, residual deviance = 9.76)
over the simpler model without this level (df=32, residual deviance =
12.39, AIC=156.63).

This latter analysis suggests that our data does not provide strong
evidence for the maintenance of belief in local readings in our
setting. But we need to be careful. So far we have not taken into
account the sequential structure of our task. It might be that local
readings are infrequent because subjects tend to exit early in the
sequence, rather than late. This is why we need preference-related
controls.

\paragraph{Preference-related controls.} Positional answers for our
preference-related control conditions are shown in
Figure~\ref{Fig:TF_judgments}.\fn{We removed 19 data points for the
  \ec-conditions because of a coding mistake that presented subjects
  with incongruent sentence-picture pairs. Including these and
  classifying them as \emph{errors} does not change the results of the
  reported analyses.}
%
\begin{figure}[]
\centering

@@

\caption{Answers for \lc-sequences and \ec-sequences. \emph{True}
  judgements are plotted upwards for each step in the sequence,
  \emph{false} judgements downwards.}
\label{Fig:TF_judgments}
\end{figure}
%
Judgments were coded as \ec or as \lc if they were as expected under
one of these readings, and as \emph{error} if they were not. The
distributions of readings thus coded is given in
Table~\ref{tab:CC-ContingencyTable-TF}. Some subjects saw 4 trials of
some target-related control conditions, but the table and the
subsequent analysis of the count data only looks at the data from the
first three encounters. This is because otherwise we would see a
spurious main effect of \textsc{Position} that is just needed to
account for the less frequent occurrence of judgements at position 4.
%
\begin{table}[]
\centering

\begin{tabular}{lcccc}
  condition & position & \multicolumn{3}{c}{response type} \\ \cmidrule(r){3-5}
  && \emph{LC} & \emph{EC} & \emph{error}  \\
  \midrule \addlinespace[0.15cm]
  \textbf{\lc neutral} 
  & Pos 1 & 27 &  7  & 4 \\
  & Pos 2 & 23 &  11 & 4 \\
  & Pos 3 & 26 &  9  & 3 \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{76} &  \textbf{27} & \textbf{1}  \\
  \addlinespace[0.25cm]
  \textbf{\lc with \lc cue} 
  & Pos 1 & 26 &  7 & 5  \\
  & Pos 2 & 31 &  5 & 2  \\
  & Pos 3 & 33 &  3 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{90} &  \textbf{15} & \textbf{9}  \\
  \addlinespace[0.25cm]
  \textbf{\lc with \ec cue} 
  & Pos 1 & 14 &  17 & 7  \\
  & Pos 2 & 15 &  16 & 7  \\
  & Pos 3 & 15 &  17 & 6  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{44} &  \textbf{50} & \textbf{20}  \\
  \addlinespace[0.25cm]
  \textbf{\ec neutral} 
  & Pos 1 & 22 &  6 & 5  \\
  & Pos 2 & 28 &  6 & 4  \\
  & Pos 3 & 29 &  7 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{79} &  \textbf{19} & \textbf{11}  \\
  \addlinespace[0.25cm]
  \textbf{\ec with \lc cue} 
  & Pos 1 & 30 &  5 & 1  \\
  & Pos 2 & 32 &  5 & 0  \\
  & Pos 3 & 31 &  3 & 2  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{93} &  \textbf{13} & \textbf{3}  \\
  \addlinespace[0.25cm]
  \textbf{\ec with \ec cue} 
  & Pos 1 & 19 &  10 & 7  \\
  & Pos 2 & 19 &  7  & 9  \\
  & Pos 3 & 19 &  7  & 11  \\ \addlinespace[0.15cm]
  & \textbf{Total} & \textbf{57} &  \textbf{24} & \textbf{27}  \\
\end{tabular}

@@

As expected, most answers classify as \lc-readings, fewer as
\ec-readings. Moreover, intonation does seem to have the expected
effect as well: as compared to neutral intonation, \lc-cueing
intonation ups the count for \lc-readings, while \ec-cueing ups the
count for \ec-readings. If we compare the number of \lc- or
\ec-readings for each intonation type across sequences, we see that
there are always more answers indicative of the relevant readings when
that reading can be judged later in the sequence.

There are only very few answers that appear to be random
mistakes. Some answers in the `error' category are plausibly
`spill-over' errors. But on top of that, a substantial number of
`error' answers occur on critical positions. In the \lc-sequence there
are 23 \emph{true} answers at step 6 where \emph{false} answers would
be taken as indicative of \ec-readings. In the \ec-sequence there are
28 \emph{false} answers at step 5 where \emph{true} answers would be
taken as indicative of \lc-readings. \todo{A general tendency to
  answer later in the sequence does explain this. We don't see a
  similar effect in the critical conditions because the literal
  reading is so dominant.} Finally, intonation patterns that cue
\ec-readings seem to have led to more mistakes than other intonation
patterns.

\begin{itemize}
\item \textcolor{red}{LOG-LINEAR MODELS}
\end{itemize}
